Eleven years of student replication projects provide evidence on the correlates of replicability in psychology

### 2023-10-16 paper/manuscript.tex
rerun models; add closeness and pre-reg stuff to ms

These replications were all conducted on short time scales, within a constrained class budget. In some cases the budget limited the number of participants who could be recruited, occasionally below what the original study included or what power analyses suggested. The replications had a median post-exclusion sample size that was 86\% of the original sample size. In nearly all cases, replications were conducted online, with recruitment from Amazon Mechanical Turk (the default from 2011 to 2020) or Prolific (the default from 2021 -- 2022).

Many different measures can be used to define replication success of an individual statistical result (\protect\hyperlink{ref-simonsohn2015}{27}--\protect\hyperlink{ref-mathur2020}{29}). However, whether a replication should be considered successful is not always dependent on only one statistical comparison between the two studies. Often in original papers, multiple statistical tests are cited in support of the claim that a pattern of results matches a particular theoretical expectation.

@@

As a complement to our primary subjective outcome, we also used two statistical measures of replication on the subset of the data where they were computable for the key statistic of interest (136 cases, see Figure \ref{fig:prisma}). We used \emph{p-original}, the \emph{p}-value on the null hypothesis that the original and replication statistics are from the same distribution, and \emph{prediction interval}, a binary measure of whether the replication statistic fell within the prediction interval of the original statistic (\protect\hyperlink{ref-mathur2020}{29}). The prediction interval depends on the level of evidence of the original study; if the effect was marginal, the prediction interval could overlap zero; thus, a replication might fall within the predictive interval, and be consistent with the original outcome, but not provide compelling evidence for the claimed effect. Conversely, large original effects with precise point estimates may have prediction intervals that do not overlap a smaller replication effect size, and thus would be inconsistent with the original outcome, even though researcher intuition might classify it as a success. Thus, these two statistical metrics each quantify the similarity between a key statistic in the original study and the replication, but they will not always match researcher intuitions on whether a study replicated.

@@

Across the 176 studies, the average subjective replication score was 49\%, which we can interpret as an overall subjective replication rate. 46\% (62/136) of replications had outcomes with point-estimates within the prediction interval of the original outcome. The median \emph{p}-original value on the original and replication point-estimates coming from the same distribution was 0.03, representing the median probability that a replication study's estimate would be at least as extreme as was actually observed, if in fact the replication and original were statistically consistent. The subjective replication scores were moderately correlated with both prediction interval (r=0.37) and \emph{p}-original (r=0.29) values; prediction interval and \emph{p}-original were highly correlated with each other (r=0.73).

@@

On average, there was a diminution of effect sizes from original to replication. This pattern of results is consistent with the results of RP:P (\protect\hyperlink{ref-openscienceconsortium2015}{3}). Specifically, the median original effect size (in SMD units) was 0.61 (interquartile range: 0.46 - 0.98), and the median replication effect size was 0.28 (0.09 - 0.76), for a median difference of 0.3 (0.03 - 0.64). Among the pairs with a subjective replication score of 1 (N=38 with standardized effect sizes), the median original effect size was 0.92 (interquartile range: 0.52 - 1.4), and the median replication effect size was 0.89 (0.57 - 1.24), for a median difference of 0.02 (-0.16 - 0.3). Thus, for the highest subjectively-rated replications, effect sizes for original and replication were similar in magnitude on average, with some replications having larger effects than their originals.

@@

\caption{\label{tab:unnamed-chunk-4}Descriptive properties of the replication dataset. For categorical features, counts and percentages are given. For continuous features, the median value and interquartile range are given. N=176 except for original effect size and p-value where N=112. }
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lc}
\toprule
Feature & Summary\\
\midrule
Subfield & \\
\hspace{1em}Cognitive & 62 (35\%)\\
\hspace{1em}Social & 70 (40\%)\\
\hspace{1em}Other psych & 24 (14\%)\\
\hspace{1em}Non-psych & 20 (11\%)\\
Open data & 52 (30\%)\\
Open materials & 83 (47\%)\\
Switch from in-person to online & 94 (53\%)\\
Original authors at Stanford & 16 (9.1\%)\\
Within subjects design & 80 (45\%)\\
Single vignette & 78 (44\%)\\
Number of trials & 6 (1, 60)\\
Publication Year & 2015 (2011, 2017)\\
Original sample size & 101 (40, 181)\\
Replication sample size & 59 (31, 125)\\
Ratio of replication/original sample sizes & 0.86 (0.41, 1.05)\\
Original effect size (SMD) & 0.61 (0.46, 0.98)\\
Original p-value & 0.0001 (0.0000, 0.0069)\\
\bottomrule
\end{tabular}
\end{table}

Some multi-site replication projects have found heterogeneity in effect sizes across replication sites (\protect\hyperlink{ref-klein2018}{7},\protect\hyperlink{ref-ebersole2020}{19},\protect\hyperlink{ref-olsson2020}{30}). As a test of sensitivity to heterogeneity, we assumed that the level of heterogeneity in hypothetical multi-site replications of our sampled articles was the same as the average level heterogeneity found by (\protect\hyperlink{ref-olsson2020}{30}) in prior multi-site replications in psychology (\(\tau=.21\) in units of standardized mean difference). Under this assumption, 64\% (72/112) of replication effect sizes are distributionally consistent with the original effect size. However, more work on understanding heterogeneity is needed to understand what levels of heterogeneity to expect across different implementations of the same experiment, and how considerations of heterogeneity should impact interpretations of both novel results and replication results.

@@

Studies with open data, which almost always also had open materials as well, tended to replicate more than studies without open data. Nearly all replication studies were conducted online, but original studies were split between using in-person and online recruitment. Replications that switched to online were less likely to replicate than those that had the same modality as the original (generally both online, in a few cases both in-person). While online studies in general show comparable results to studies conducted in person (\protect\hyperlink{ref-crump2013}{31}), switching the modality does decrease the closeness of the replication, and some studies done in person may not have been well-adapted (e.g., inductions might have been weaker or instructions might have been insufficient). These factors -- open materials, open data, and online samples -- may have all contributed to the closeness of the replications. However, they are also practices that have increased over time, and so these effects may partially reflect temporal trends. Nonetheless, publication year only very weakly correlated with replicability, but does correlate strongly with online samples and openness (Figure \ref{fig:corr-plot}).

@@

While a number of features show individual correlations with the subjective replication score, many correlate with one another. To determine which predictors were the strongest, we modeled subjective replication score as an ordinal outcome using ordinal Bayesian regression models with a logit link function, regularized using a horseshoe shrinkage prior (\protect\hyperlink{ref-carvalho09}{32}). This model estimates odds ratios representing the association of the predictor with having a higher versus lower subjective replication score. We first ran models using all of the original-replication pairs (N=176), but without original effect size and original \(p\) value as predictors, as they were uncodable for some pairs. We next ran models including all predictors, but on only the subset of data where all predictors were available (N=112).

@@

Within-participants designs stand out as the strongest correlate of replicability in the model with all the data (OR = 2.56, 95\% CrI = {[}1, 7.36{]}). In the model with all predictors, but less data, within-participants designs remain strong (OR = 2.84, 95\% CrI = {[}0.95, 10.9{]}), and standardized effect size is also substantially related to subjective replication score (OR = 3.32, 95\% CrI = {[}1.04, 11.99{]}). Both effects are robust to a sensitivity analysis including only studies with close replications and matching statistical tests (within-participants OR = 3.7, 95\% CrI = {[}0.98, 21.2{]}; effect size OR = 7.34, 95\% CrI = {[}1.25, 48.65{]}).

We also ran models predicting our secondary outcome measures: a logistic model predicting whether the replication effect was within the prediction interval of the original effect, and a linear model predicting what the \(p\)-original was between the replication and original. Both these models had more uncertain estimates. While the credible intervals were wide, the general patterns of predictor direction and relative strength were similar to the subjective replication models (see Supplement for results from all models). The strongest predictors for prediction interval were still within-participants designs (OR = 1.89, 95\% CrI = {[}0.84, 8.98{]}) and studies with larger effect sizes (OR = 1.2, 95\% CrI = {[}0.75, 2.81{]}).

@@

Our results are limited by the sample of studies we included, which are limited in number and may not be representative of the studies of interest to psychologists as a whole. Further, our predictor variables were not manipulated, so they cannot be interpreted as causing (non-)replication, but only as correlational markers. Some of the correlates are most easily interpreted as being about the original study, and others reflect the closeness of the replication to the original. For instance, while within-participants designs are more likely to replicate than between-participants designs, this predictor could also be related to the types of experiments that tend to be run in each design. Given the strong relationship of this factor to replicability, slightly more skepticism and critical reading of between-participants designs may be warranted, but this correlation, by itself, does not mean scientists should prefer within-participants designs.

Large scale replication studies are costly and arduous. The batch of replications presented here were pedagogical replications, done as part of a class. Trainees must learn experimental methods, and conducting replications as part of methods classes serves several purposes: it enables students to learn to do experiments in a supportive context, it often leads to more useful results than if students designed their own experiments from scratch, and it creates a resource for studying the literature (\protect\hyperlink{ref-quintana2021}{33}--\protect\hyperlink{ref-hawkins2018}{36}). We believe that this kind of pedagogy has an important role to play in improving methodological practices in psychology more broadly. The tools and workflows of rigorous, replicable science cannot simply be mandated: they have to be learned.

@@

Our pre-registration, code, and coded data are available at \url{https://osf.io/xwn9m/} (\protect\hyperlink{ref-osfdata}{37}).

@@

The dataset of replication projects comes from class projects conducted in PSYCH 251 (earlier called PSYCH 254) a graduate-level experimental methods class taught at Stanford by MCF from 2011 to 2022. This class is commonly taken by first year graduate students in psychology and related disciplines, and it has been a requirement of the Psychology PhD since around 2015. Each student chose a study to replicate, implemented the study, wrote analysis code, pre-registered their replication, ran the study, and turned in a structured final report including methods, analytic plan, changes from the original study, confirmatory and exploratory analyses, and discussion of outcomes. Students were encouraged to do experimental replications, but some students chose to replicate correlational outcomes or do computational reproducibility projects instead. We cannot include the full student reports for confidentiality reasons, but we include over 50 reports that we received permission to share and the template given to students at \url{https://osf.io/xwn9m/} (\protect\hyperlink{ref-osfdata}{37}).

@@

Four of the replication projects were included in RP:P, and 10 were previously reported in (\protect\hyperlink{ref-hawkins2018}{36}) (which reported 11 student replications from the 2015-2016 class; one of those was excluded from the current sample for being non-experimental).

@@

This subjective replication outcome was chosen because it already existed, could be applied to all projects (regardless of type and detail of statistical reporting), and did not rely solely on one statistical measure. As a complement, we also identified a ``key'' statistical test for each paper (see below for details), and if possible, computed \(p\)-original and prediction interval at this statistic, following (\protect\hyperlink{ref-mathur2020}{29}). \(p\)-original was a continuous measure of the \(p\) value on the hypothesis that the original and replication samples come from the same distribution. Prediction interval was a binary measure of whether the replication outcome fell within the prediction interval of the original outcome measure.

@@

We used a horseshoe shrinkage prior on the fixed effect coefficients because we had a lot of predictors compared to the amount of data (\protect\hyperlink{ref-carvalho09}{32}). All models included random slopes for predictors nested within year the class occurred to control for variation between cohorts of students. We did not include any interaction terms in the models. All numeric predictor variables were z-scored after other transforms (e.g., logs) to ensure comparable regularization effects from the horseshoe prior. The priors we used were horseshoe(3) for betas, normal(0,.5) for standard deviation of random slopes, and lkj(1) for correlations between random slopes. Models were run in BRMS (\protect\hyperlink{ref-brms}{38}).

@@

\leavevmode\vadjust pre{\hypertarget{ref-simonsohn2015}{}}%
\CSLLeftMargin{27. }%
\CSLRightInline{Simonsohn U. Small {Telescopes}: {Detectability} and the {Evaluation} of {Replication Results}. Psychol Sci {[}Internet{]}. 2015 May 1 {[}cited 2023 Apr 4{]};26(5):559--69. Available from: \url{https://doi.org/10.1177/0956797614567341}}

\leavevmode\vadjust pre{\hypertarget{ref-gelman2018}{}}%
\CSLLeftMargin{28. }%
\CSLRightInline{Gelman A. Don't characterize replications as successes or failures. Behav Brain Sci {[}Internet{]}. 2018/ed {[}cited 2023 Apr 4{]};41:e128. Available from: \url{https://www.cambridge.org/core/journals/behavioral-and-brain-sciences/article/abs/dont-characterize-replications-as-successes-or-failures/1AB661F1E7A08E9C5870258878DBA0EA}}

\leavevmode\vadjust pre{\hypertarget{ref-mathur2020}{}}%
\CSLLeftMargin{29. }%
\CSLRightInline{Mathur MB, VanderWeele TJ. New statistical metrics for multisite replication projects. J R Stat Soc Ser A Stat Soc {[}Internet{]}. 2020 {[}cited 2023 Apr 4{]};183(3):1145--66. Available from: \url{https://onlinelibrary.wiley.com/doi/abs/10.1111/rssa.12572}}

\leavevmode\vadjust pre{\hypertarget{ref-olsson2020}{}}%
\CSLLeftMargin{30. }%
\CSLRightInline{Olsson-Collentine A, Assen MA van, Wicherts J. Heterogeneity in direct replications in psychology and its association with effect size. 2020; }

\leavevmode\vadjust pre{\hypertarget{ref-crump2013}{}}%
\CSLLeftMargin{31. }%
\CSLRightInline{Crump MJC, McDonnell JV, Gureckis TM. Evaluating {Amazon}'s {Mechanical Turk} as a {Tool} for {Experimental Behavioral Research}. PLOS ONE {[}Internet{]}. 2013 Mar 13 {[}cited 2020 Sep 30{]};8(3):e57410. Available from: \url{https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0057410}}

\leavevmode\vadjust pre{\hypertarget{ref-carvalho09}{}}%
\CSLLeftMargin{32. }%
\CSLRightInline{Carvalho CM, Polson NG, Scott JG. Handling sparsity via the horseshoe. In: Dyk D van, Welling M, editors. Proceedings of the twelth international conference on artificial intelligence and statistics {[}Internet{]}. Hilton Clearwater Beach Resort, Clearwater Beach, Florida USA: PMLR; 2009. p. 73--80. (Proceedings of machine learning research; vol. 5). Available from: \url{https://proceedings.mlr.press/v5/carvalho09a.html}}

\leavevmode\vadjust pre{\hypertarget{ref-quintana2021}{}}%
\CSLLeftMargin{33. }%
\CSLRightInline{Quintana DS. Replication studies for undergraduate theses to improve science and education. Nat Hum Behav {[}Internet{]}. 2021 Sep {[}cited 2023 Apr 4{]};5(9, 9):1117--8. Available from: \url{https://www.nature.com/articles/s41562-021-01192-8}}

\leavevmode\vadjust pre{\hypertarget{ref-wagge2019}{}}%
\CSLLeftMargin{34. }%
\CSLRightInline{Wagge JR, Brandt MJ, Lazarevic LB, Legate N, Christopherson C, Wiggins B, et al. Publishing {Research With Undergraduate Students} via {Replication Work}: {The Collaborative Replications} and {Education Project}. Front Psychol {[}Internet{]}. 2019 {[}cited 2023 Apr 4{]};10. Available from: \url{https://www.frontiersin.org/articles/10.3389/fpsyg.2019.00247}}

\leavevmode\vadjust pre{\hypertarget{ref-frank2012}{}}%
\CSLLeftMargin{35. }%
\CSLRightInline{Frank MC, Saxe R. Teaching {Replication}: Perspect Psychol Sci {[}Internet{]}. 2012 Nov 7 {[}cited 2020 Sep 14{]}; Available from: \url{https://journals.sagepub.com/doi/10.1177/1745691612460686}}

\leavevmode\vadjust pre{\hypertarget{ref-hawkins2018}{}}%
\CSLLeftMargin{36. }%
\CSLRightInline{Hawkins RD, Smith EN, Au C, Arias JM, Catapano R, Hermann E, et al. Improving the replicability of psychological science through pedagogy. Advances in Methods and Practices in Psychological Science. 2018 Mar;1(1):7--18. }

\leavevmode\vadjust pre{\hypertarget{ref-osfdata}{}}%
\CSLLeftMargin{37. }%
\CSLRightInline{Boyce V. Meta analysis of class replication projects from PSYCH251 {[}Internet{]}. OSF; 2023. Available from: \href{https://osf.io/xwn9m}{osf.io/xwn9m}}

>>>>

These replications were all conducted on short time scales, within a constrained class budget. In some cases the budget limited the number of participants who could be recruited, occasionally below what the original study included or what power analyses suggested. The replications had a median post-exclusion sample size that was 86\% of the original sample size. In nearly all cases, replications were conducted online, with recruitment from Amazon Mechanical Turk (the default from 2011 to 2020) or Prolific (the default from 2021 -- 2022). The goal was for replications to be as close as possible to the original, but budgetary constraints, inability to option original materials, and primarily using online samples meant that replications varied in their degree of closeness to the original. According to the schema from (\protect\hyperlink{ref-lebel2018}{27}), 19\% of studies were exact replications, 29\% were very close replications, 44\% were close replications, and 8 were far replications (See Methods for more details on common deviations).

Many different measures can be used to define replication success of an individual statistical result (\protect\hyperlink{ref-simonsohn2015}{28}--\protect\hyperlink{ref-mathur2020}{30}). However, whether a replication should be considered successful is not always dependent on only one statistical comparison between the two studies. Often in original papers, multiple statistical tests are cited in support of the claim that a pattern of results matches a particular theoretical expectation.

@@

As a complement to our primary subjective outcome, we also used two statistical measures of replication on the subset of the data where they were computable for the key statistic of interest (136 cases, see Figure \ref{fig:prisma}). We used \emph{p-original}, the \emph{p}-value on the null hypothesis that the original and replication statistics are from the same distribution, and \emph{prediction interval}, a binary measure of whether the replication statistic fell within the prediction interval of the original statistic (\protect\hyperlink{ref-mathur2020}{30}). The prediction interval depends on the level of evidence of the original study; if the effect was marginal, the prediction interval could overlap zero; thus, a replication might fall within the predictive interval, and be consistent with the original outcome, but not provide compelling evidence for the claimed effect. Conversely, large original effects with precise point estimates may have prediction intervals that do not overlap a smaller replication effect size, and thus would be inconsistent with the original outcome, even though researcher intuition might classify it as a success. Thus, these two statistical metrics each quantify the similarity between a key statistic in the original study and the replication, but they will not always match researcher intuitions on whether a study replicated.

@@

Across the 176 studies, the average subjective replication score was 49\%, which we can interpret as an overall subjective replication rate. 45\% (61/136) of replications had outcomes with point-estimates within the prediction interval of the original outcome. The median \emph{p}-original value on the original and replication point-estimates coming from the same distribution was 0.03, representing the median probability that a replication study's estimate would be at least as extreme as was actually observed, if in fact the replication and original were statistically consistent. The subjective replication scores were moderately correlated with both prediction interval (r=0.36) and \emph{p}-original (r=0.27) values; prediction interval and \emph{p}-original were highly correlated with each other (r=0.73).

@@

On average, there was a diminution of effect sizes from original to replication. This pattern of results is consistent with the results of RP:P (\protect\hyperlink{ref-openscienceconsortium2015}{3}). Specifically, the median original effect size (in SMD units) was 0.61 (interquartile range: 0.46 - 0.98), and the median replication effect size was 0.28 (0.09 - 0.76), for a median difference of 0.31 (0.04 - 0.66). Among the pairs with a subjective replication score of 1 (N=38 with standardized effect sizes), the median original effect size was 0.92 (interquartile range: 0.52 - 1.4), and the median replication effect size was 0.89 (0.57 - 1.24), for a median difference of 0.03 (-0.16 - 0.33). Thus, for the highest subjectively-rated replications, effect sizes for original and replication were similar in magnitude on average, with some replications having larger effects than their originals.

@@

\caption{\label{tab:unnamed-chunk-5}Descriptive properties of the replication dataset. For categorical features, counts and percentages are given. For continuous features, the median value and interquartile range are given. N=176 except for original effect size and p-value where N=112. }
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lc}
\toprule
Feature & Summary\\
\midrule
Subfield & \\
\hspace{1em}Cognitive & 62 (35\%)\\
\hspace{1em}Social & 70 (40\%)\\
\hspace{1em}Other psych & 24 (14\%)\\
\hspace{1em}Non-psych & 20 (11\%)\\
Open data & 52 (30\%)\\
Open materials & 83 (47\%)\\
Switch from in-person to online & 94 (53\%)\\
Original authors at Stanford & 16 (9.1\%)\\
Within participants design & 80 (45\%)\\
Single vignette & 78 (44\%)\\
Number of trials & 6 (1, 60)\\
Publication Year & 2015 (2011, 2017)\\
Original sample size & 101 (40, 181)\\
Replication sample size & 59 (31, 125)\\
Ratio of replication/original sample sizes & 0.86 (0.41, 1.05)\\
Original effect size (SMD) & 0.61 (0.46, 0.98)\\
Original p-value & 0.0001 (0.0000, 0.0069)\\
\bottomrule
\end{tabular}
\end{table}

Some multi-site replication projects have found heterogeneity in effect sizes across replication sites (\protect\hyperlink{ref-klein2018}{7},\protect\hyperlink{ref-ebersole2020}{19},\protect\hyperlink{ref-olsson2020}{31}). As a test of sensitivity to heterogeneity, we assumed that the level of heterogeneity in hypothetical multi-site replications of our sampled articles was the same as the average level heterogeneity found by (\protect\hyperlink{ref-olsson2020}{31}) in prior multi-site replications in psychology (\(\tau=.21\) in units of standardized mean difference). Under this assumption, 63\% (71/112) of replication effect sizes are distributionally consistent with the original effect size. However, more work on understanding heterogeneity is needed to understand what levels of heterogeneity to expect across different implementations of the same experiment, and how considerations of heterogeneity should impact interpretations of both novel results and replication results.

@@

Studies with open data, which almost always also had open materials as well, tended to replicate more than studies without open data. Nearly all replication studies were conducted online, but original studies were split between using in-person and online recruitment. Replications that switched to online were less likely to replicate than those that had the same modality as the original (generally both online, in a few cases both in-person). While online studies in general show comparable results to studies conducted in person (\protect\hyperlink{ref-crump2013}{32}), switching the modality does decrease the closeness of the replication, and some studies done in person may not have been well-adapted (e.g., inductions might have been weaker or instructions might have been insufficient). These factors -- open materials, open data, and online samples -- may have all contributed to the closeness of the replications. However, they are also practices that have increased over time, and so these effects may partially reflect temporal trends. Nonetheless, publication year only very weakly correlated with replicability, but does correlate strongly with online samples and openness (Figure \ref{fig:corr-plot}).

@@

While a number of features show individual correlations with the subjective replication score, many correlate with one another. To determine which predictors were the strongest, we modeled subjective replication score as an ordinal outcome using ordinal Bayesian regression models with a logit link function, regularized using a horseshoe shrinkage prior (\protect\hyperlink{ref-carvalho09}{33}). This model estimates odds ratios representing the association of the predictor with having a higher versus lower subjective replication score. We first ran models using all of the original-replication pairs (N=176), but without original effect size and original \(p\) value as predictors, as they were uncodable for some pairs. We next ran models including all predictors, but on only the subset of data where all predictors were available (N=112).

@@

Within-participants designs stand out as the strongest correlate of replicability in the model with all the data (OR = 2.59, 95\% CrI = {[}1, 7.3{]}). In the model with all predictors, but less data, within-participants designs remain strong (OR = 2.89, 95\% CrI = {[}0.96, 11.24{]}), and standardized effect size is also substantially related to subjective replication score (OR = 3.33, 95\% CrI = {[}1.11, 12.44{]}). Both effects are robust to a sensitivity analysis including only studies with close replications and matching statistical tests (within-participants OR = 3.58, 95\% CrI = {[}0.91, 19.57{]}; effect size OR = 7.22, 95\% CrI = {[}1.24, 47{]}).

We also ran models predicting our secondary outcome measures: a logistic model predicting whether the replication effect was within the prediction interval of the original effect, and a linear model predicting what the \(p\)-original was between the replication and original. Both these models had more uncertain estimates. While the credible intervals were wide, the general patterns of predictor direction and relative strength were similar to the subjective replication models (see Supplement for results from all models). The strongest predictors for prediction interval were still within-participants designs (OR = 1.78, 95\% CrI = {[}0.86, 8.3{]}) and studies with larger effect sizes (OR = 0.98, 95\% CrI = {[}0.56, 1.63{]}).

@@

Our results are limited by the sample of studies we included, which are limited in number and may not be representative of the studies of interest to psychologists as a whole. Further, our predictor variables were not manipulated, so they cannot be interpreted as causing (non-)replication, but only as correlational markers. Some of the correlates are most easily interpreted as being about the original study, and others reflect the closeness of the replication to the original.

For instance, while within-participants designs are more likely to replicate than between-participants designs, this predictor could also be related to the types of experiments that tend to be run in each design. Additionally, as (\protect\hyperlink{ref-greenwald1976}{25}) notes, within-participants designs can lead to practice effects, carry-over of treatments, and critically, sensitization effects, where participants reason about the contrast between conditions and respond differently due to that reasoning. Sensitization effects are a threat to internal validity, and could be replicable even in the absence of the effect itself being replicable. Given the strong relationship of within-participants designs to replicability, slightly more skepticism and critical reading of between-participants designs may be warranted, but this correlation, by itself, does not mean scientists should prefer within-participants designs.

Large scale replication studies are costly and arduous. The batch of replications presented here were pedagogical replications, done as part of a class. Trainees must learn experimental methods, and conducting replications as part of methods classes serves several purposes: it enables students to learn to do experiments in a supportive context, it often leads to more useful results than if students designed their own experiments from scratch, and it creates a resource for studying the literature (\protect\hyperlink{ref-quintana2021}{34}--\protect\hyperlink{ref-hawkins2018}{37}). We believe that this kind of pedagogy has an important role to play in improving methodological practices in psychology more broadly. The tools and workflows of rigorous, replicable science cannot simply be mandated: they have to be learned.

@@

Our pre-registration, code, and coded data are available at \url{https://osf.io/xwn9m/} (\protect\hyperlink{ref-osfdata}{38}).

At the onset of this project, all authors had some familiarity with the dataset. MCF taught PSYCH 254/251; MM and VB had each been students in the class. The coding of the data primarily took place before the pre-registration was finalized; the only way for us to know what variables were actually codeable and what forms of missingness were present in the data was to code it. Thus, the pre-registration was a pre-registration of the analyses and a ``locking-in'' of the coding scheme. VB coded both the predictor variables and was one of the coders of outcome variables, but these were coded in separate passes through the data.

\hypertarget{deviations-from-pre-registration}{%
\subsection{Deviations from pre-registration}\label{deviations-from-pre-registration}}

The descriptive statistics and bivariate correlations between predictor and outcome variables were not pre-registered.

In the pre-registration, our research questions mentioned publication journal, but due to the distribution of journals, journal was not a predictor in the pre-registered model plan (or the actual models). Our pre-registration stated that we could not include students names or reports in the released dataset; we ended up letting students opt-in to associating their names with their projects and having their replication report included. Our pre-registration stated that any errors found would be noted; errors are noted in the data spreadsheet of the materials.

@@

The dataset of replication projects comes from class projects conducted in PSYCH 251 (earlier called PSYCH 254) a graduate-level experimental methods class taught at Stanford by MCF from 2011 to 2022. This class is commonly taken by first year graduate students in psychology and related disciplines, and it has been a requirement of the Psychology PhD since around 2015. Each student chose a study to replicate, implemented the study, wrote analysis code, pre-registered their replication, ran the study, and turned in a structured final report including methods, analytic plan, changes from the original study, confirmatory and exploratory analyses, and discussion of outcomes. Students were encouraged to do experimental replications, but some students chose to replicate correlational outcomes or do computational reproducibility projects instead. We cannot include the full student reports for confidentiality reasons, but we include over 50 reports that we received permission to share and the template given to students at \url{https://osf.io/xwn9m/} (\protect\hyperlink{ref-osfdata}{38}).

@@

Replications varied in how close they were to the original; while the goal was to replicate the original as closely as possible, some deviations were sometimes necessary. Student reports contained a section listing changes from the original. In response to a reviewer, we attempted to code studies using the classification scheme from (\protect\hyperlink{ref-lebel2018}{27}).

Common reasons for replications being very close instead of exact were a switch from in-person to online and lack of access to the original instructions. (It was not always possible to tell if students had access to original wording and presentation style of materials.) Common reason for replications being only close included recreating materials when the original materials were not available and changing materials to fit the audience (ex. switching from UK to US English). (It was not always possible to determine if students had the original materials; in some cases students mentioned trying to obtain materials but did not indicate if they had succeeded.) Less common reasons for replications being only close included changing the number of trials per participant or reducing training periods.

Far replications were rare, but were primarily due to a couple deviations. In some cases an original study had a specific population (ex. high schoolers or hiring managers), and the replication was on a convenience population. The other main reason was changing the language of materials to English when cultural or linguistic factors were potentially relevant to the construct of interest.

Other common changes (that we did not consider as decreasing the closeness) were switching which convenience population was used (i.e.~from college students to crowdworkers or from one crowdwork platform to another), changing or reducing demographic surveys (which were used only in exploratory or secondary analyses, if at all), and removing secondary measures that were not part of the analyses being replicated.
Four of the replication projects were included in RP:P, and 10 were previously reported in (\protect\hyperlink{ref-hawkins2018}{37}) (which reported 11 student replications from the 2015-2016 class; one of those was excluded from the current sample for being non-experimental).

@@

This subjective replication outcome was chosen because it already existed, could be applied to all projects (regardless of type and detail of statistical reporting), and did not rely solely on one statistical measure. As a complement, we also identified a ``key'' statistical test for each paper (see below for details), and if possible, computed \(p\)-original and prediction interval at this statistic, following (\protect\hyperlink{ref-mathur2020}{30}). \(p\)-original was a continuous measure of the \(p\) value on the hypothesis that the original and replication samples come from the same distribution. Prediction interval was a binary measure of whether the replication outcome fell within the prediction interval of the original outcome measure.

@@

We used a horseshoe shrinkage prior on the fixed effect coefficients because we had a lot of predictors compared to the amount of data (\protect\hyperlink{ref-carvalho09}{33}). All models included random slopes for predictors nested within year the class occurred to control for variation between cohorts of students. We did not include any interaction terms in the models. All numeric predictor variables were z-scored after other transforms (e.g., logs) to ensure comparable regularization effects from the horseshoe prior. The priors we used were horseshoe(3) for betas, normal(0,.5) for standard deviation of random slopes, and lkj(1) for correlations between random slopes. Models were run in BRMS (\protect\hyperlink{ref-brms}{39}).

@@

\leavevmode\vadjust pre{\hypertarget{ref-lebel2018}{}}%
\CSLLeftMargin{27. }%
\CSLRightInline{LeBel EP, McCarthy RJ, Earp BD, Elson M, Vanpaemel W. A {Unified Framework} to {Quantify} the {Credibility} of {Scientific Findings}. }

\leavevmode\vadjust pre{\hypertarget{ref-simonsohn2015}{}}%
\CSLLeftMargin{28. }%
\CSLRightInline{Simonsohn U. Small {Telescopes}: {Detectability} and the {Evaluation} of {Replication Results}. Psychol Sci {[}Internet{]}. 2015 May 1 {[}cited 2023 Apr 4{]};26(5):559--69. Available from: \url{https://doi.org/10.1177/0956797614567341}}

\leavevmode\vadjust pre{\hypertarget{ref-gelman2018}{}}%
\CSLLeftMargin{29. }%
\CSLRightInline{Gelman A. Don't characterize replications as successes or failures. Behav Brain Sci {[}Internet{]}. 2018/ed {[}cited 2023 Apr 4{]};41:e128. Available from: \url{https://www.cambridge.org/core/journals/behavioral-and-brain-sciences/article/abs/dont-characterize-replications-as-successes-or-failures/1AB661F1E7A08E9C5870258878DBA0EA}}

\leavevmode\vadjust pre{\hypertarget{ref-mathur2020}{}}%
\CSLLeftMargin{30. }%
\CSLRightInline{Mathur MB, VanderWeele TJ. New statistical metrics for multisite replication projects. J R Stat Soc Ser A Stat Soc {[}Internet{]}. 2020 {[}cited 2023 Apr 4{]};183(3):1145--66. Available from: \url{https://onlinelibrary.wiley.com/doi/abs/10.1111/rssa.12572}}

\leavevmode\vadjust pre{\hypertarget{ref-olsson2020}{}}%
\CSLLeftMargin{31. }%
\CSLRightInline{Olsson-Collentine A, Assen MA van, Wicherts J. Heterogeneity in direct replications in psychology and its association with effect size. 2020; }

\leavevmode\vadjust pre{\hypertarget{ref-crump2013}{}}%
\CSLLeftMargin{32. }%
\CSLRightInline{Crump MJC, McDonnell JV, Gureckis TM. Evaluating {Amazon}'s {Mechanical Turk} as a {Tool} for {Experimental Behavioral Research}. PLOS ONE {[}Internet{]}. 2013 Mar 13 {[}cited 2020 Sep 30{]};8(3):e57410. Available from: \url{https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0057410}}

\leavevmode\vadjust pre{\hypertarget{ref-carvalho09}{}}%
\CSLLeftMargin{33. }%
\CSLRightInline{Carvalho CM, Polson NG, Scott JG. Handling sparsity via the horseshoe. In: Dyk D van, Welling M, editors. Proceedings of the twelth international conference on artificial intelligence and statistics {[}Internet{]}. Hilton Clearwater Beach Resort, Clearwater Beach, Florida USA: PMLR; 2009. p. 73--80. (Proceedings of machine learning research; vol. 5). Available from: \url{https://proceedings.mlr.press/v5/carvalho09a.html}}

\leavevmode\vadjust pre{\hypertarget{ref-quintana2021}{}}%
\CSLLeftMargin{34. }%
\CSLRightInline{Quintana DS. Replication studies for undergraduate theses to improve science and education. Nat Hum Behav {[}Internet{]}. 2021 Sep {[}cited 2023 Apr 4{]};5(9, 9):1117--8. Available from: \url{https://www.nature.com/articles/s41562-021-01192-8}}

\leavevmode\vadjust pre{\hypertarget{ref-wagge2019}{}}%
\CSLLeftMargin{35. }%
\CSLRightInline{Wagge JR, Brandt MJ, Lazarevic LB, Legate N, Christopherson C, Wiggins B, et al. Publishing {Research With Undergraduate Students} via {Replication Work}: {The Collaborative Replications} and {Education Project}. Front Psychol {[}Internet{]}. 2019 {[}cited 2023 Apr 4{]};10. Available from: \url{https://www.frontiersin.org/articles/10.3389/fpsyg.2019.00247}}

\leavevmode\vadjust pre{\hypertarget{ref-frank2012}{}}%
\CSLLeftMargin{36. }%
\CSLRightInline{Frank MC, Saxe R. Teaching {Replication}: Perspect Psychol Sci {[}Internet{]}. 2012 Nov 7 {[}cited 2020 Sep 14{]}; Available from: \url{https://journals.sagepub.com/doi/10.1177/1745691612460686}}

\leavevmode\vadjust pre{\hypertarget{ref-hawkins2018}{}}%
\CSLLeftMargin{37. }%
\CSLRightInline{Hawkins RD, Smith EN, Au C, Arias JM, Catapano R, Hermann E, et al. Improving the replicability of psychological science through pedagogy. Advances in Methods and Practices in Psychological Science. 2018 Mar;1(1):7--18. }

\leavevmode\vadjust pre{\hypertarget{ref-osfdata}{}}%
\CSLLeftMargin{38. }%
\CSLRightInline{Boyce V. Meta analysis of class replication projects from PSYCH251 {[}Internet{]}. OSF; 2023. Available from: \href{https://osf.io/xwn9m}{osf.io/xwn9m}}
